---
title: "aa_path_network"
author: "Kimberly Ledger"
date: "3/11/2022"
output: github_document
---

this code creates a network plot of pathogen co-infections found in Amblyomma americanum 

libraries
```{r, message = FALSE}
library(tidyverse)
library(igraph)
library(intergraph)
library(ggnetwork)
```


read in the data - this is an edge list where each row represents a connection between a microbe in the column (micro1) and another column (micro2)
```{r, message = FALSE}
edge_data <- read.csv("AA_pathogen_edgelist.csv")
```

make object with just co-infections 
```{r, message = FALSE}
links <- edge_data[,c(4,5)]
```

create network object
```{r, message = FALSE}
network <- graph_from_data_frame(d=links, directed = F)
```

plot it 
```{r, message = FALSE}
plot(network)
```

plot in a circle
```{r, message = FALSE}
plot(network, layout = layout_in_circle(network))
```

make an adjacency matrix 

```{r, message = FALSE}
adj_mat <- get.adjacency(network)
adj_mat
```



try making a prettier plot - weight the connections by coinfections 
```{r, message = FALSE}
plot(network, vertex.label.color="black", 
     vertex.label.cex=.5,vertex.label.dist=0.2, 
     edge.curved=0.5, edge.width = edge.betweenness(network),
     layout=layout_in_circle(network))

#plot(network, vertex.label.color="black", 
#     vertex.label.cex=.5,vertex.label.dist=0.2, 
#     edge.curved=0.5, edge.width = edge.betweenness(network), edge.color = edge.betweenness(network),
#     layout=layout_in_circle(network))
```

```{r}
plot(network, 
     vertex.size = 15, vertex.color = "#486090", vertex.frame.color = "black",
     vertex.label.cex=.8, vertex.label.dist= 2,
     vertex.label.color = "black",
     vertex.label.font = 4, 
     vertex.label.degree = c(-pi/2,-pi/2,-pi/2, -pi/2,-pi/2,-pi/2, pi/2, pi/2, pi/2 , pi/2),
     edge.curved=0.5, edge.width = edge.betweenness(network)*2, #edge.color = "#f0a860",
     layout=layout_in_circle(network))

```

saved using: 
tiff(filename = "coinf_grey.tiff", height = 2000, width = 2000, res = 300)
