---
title: "aa_pathogens"
author: "Kimberly Ledger"
date: "2/17/2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

# *Amblyomma americanum* pathogen analyses

This code will explore the pathogens *Amblyomma americanum* collected in Alachua Co.

Load libraries 
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(lattice)
library(lme4)
library(MuMIn)
#library(car)
#library(MASS)
#library(stargazer)
#library(effects)
```

## Part 1: Prepare the data
Read in raw Alachua County Tick Data
```{r}
ticks <- read.csv("AlachuaTicksAllSites.csv")
str(ticks)
```

Read in raw Pathogen Data 
```{r}
pathogen <- read.csv("AlCo_Pathogens.csv")
str(pathogen)
```

Join pathogen results with the raw tick data by the gDNA_Sample and Collection_Sample identities

```{r}
data <- ticks %>%
  dplyr::left_join(pathogen, by = c("gDNA_Sample", "Collection_Sample"))

#data <- ticks %>%
#  dplyr::left_join(pathogen, by = c("Collection_Sample"))

head(data)
```
for now there are NA's in the data for any tick sample that was not processed in the lab

verify the data was joined correctly....
```{r}
tc <- data %>%
  filter(Theileria_cervi== 1)
```
all good to go except for something is off with Rickettsia amblyommatis.. this data only shows 563 positive but the original spreadsheet indicates 583 positive... moving on for now but will need to fix if we want to run any analyses on R.amblyommatis later

Make a summary table of microorganisms detected in each tick species 
```{r}
data %>%
  dplyr::select(-c(Visit, Tick_No_Tube)) %>%
  group_by(Species) %>%
  dplyr::summarise(across(where(is.integer), sum, na.rm = TRUE))
```

Okay, what PATHOGENIC microorganisms to humans or domestic animals are presest in *A. americanum*

* Babesia sp Coco (8)
* Cytauxozoon felis (1)
* Ehrlichia chaffeensis (7)
* Ehrlichia ewingii (10)
* Ehrlichia sp. PM (9) 
* Rickettsia parkeri (3)


Now i want to separate out just the *Amblyomma americanum* to look at... 
```{r}
aa <- data %>%
  filter(Species == "AA") %>%
  mutate_all(na_if, "")
head(aa)
```

Read in the covariate data for all 18 sites 
This site contains information on if the environment type was manicured or natural, and the forest, PC1 and PC2 landscape buffers
```{r}
covar <- read.csv("AlCo_SitesCovariates.csv")  ## 18 sites 
head(covar)
```

create columns that indicate the presence or absence of the pathogens/pathogen groups of interest for *Amblyomma americanum*

* all human and domestic animal pathogens
* all human and domestic animal Ehrlichia sp.
* *Theileria cervi* - will not be included in this manuscript but could be interesting for something later on...

```{r}
aa_join <- aa %>%
  filter(!is.na(gDNA_Sample)) %>%
  dplyr::select(Site_ID, Babesia_sp_Coco, Babesia_odocoieli, Cytauxozoon_felis, Ehrlichia_chaffeensis, Ehrlichia_ewingii, Ehrlichia_sp_PanolaMountain, Rickettsia_parkeri, Theileria_cervi) %>%
  mutate(zoo_path = Babesia_sp_Coco + Babesia_odocoieli + Cytauxozoon_felis + Ehrlichia_chaffeensis + Ehrlichia_ewingii + Ehrlichia_sp_PanolaMountain + Rickettsia_parkeri) %>%
  mutate(zoo_ehr = Ehrlichia_chaffeensis + Ehrlichia_ewingii + Ehrlichia_sp_PanolaMountain) %>%
  dplyr::select(Site_ID, zoo_path, zoo_ehr, Theileria_cervi) %>%
  dplyr::left_join(covar, by = "Site_ID")

aa_join$zoo_path[aa_join$zoo_path > 0] <- 1  ## set coinfections to 1 
aa_join$zoo_ehr[aa_join$zoo_ehr > 0] <- 1  ## set coinfections to 1 

head(aa_join)
```

now let's plot the relationship of pres/absence of pathogens with PC1 
```{r}
#plot(x=aa_join$PC1_500m, y=aa_join$zoo_path)
xyplot(zoo_path ~ PC1_500m | Field_Type, 
       data = aa_join)

#plot(x=aa_join$PC1_500m, y=aa_join$zoo_ehr)
xyplot(zoo_ehr ~ PC1_500m | Field_Type, 
       data = aa_join)

#plot(x=aa_join$PC1_500m, y=aa_join$Theileria_cervi)
xyplot(Theileria_cervi~ PC1_500m | Field_Type, 
       data = aa_join)
```

#### model presence/absence of any zoonotic pathogen at the individual level

try out mixed-effects GLM w/ binomial distribution - using Site ID as a random effect
```{r, message=FALSE}
B0 <- glmer(zoo_path ~ 1 + (1|Site_ID), family = binomial, data = aa_join)
B1 <- glmer(zoo_path ~ Field_Type + (1|Site_ID), family = binomial, data = aa_join)
B2 <- glmer(zoo_path ~ PC1_500m + (1|Site_ID), family = binomial, data = aa_join)
B3 <- glmer(zoo_path ~ PC1_500m + Field_Type + (1|Site_ID), family = binomial, data = aa_join)
B4 <- glmer(zoo_path ~ PC1_500m * Field_Type + (1|Site_ID), family = binomial, data = aa_join)

AICc(B0,B1,B2,B3,B4)
```

null model is as good or better than others
```{r}
summary(B4)
```

#### model presence/absence of zoonotic Ehrlichia at the individual level

try out mixed-effects GLM w/ binomial distribution - using Site ID as a random effect
```{r, message=FALSE}
E0 <- glmer(zoo_ehr ~ 1 + (1|Site_ID), family = binomial, data = aa_join)
E1 <- glmer(zoo_ehr ~ Field_Type + (1|Site_ID), family = binomial, data = aa_join)
E2 <- glmer(zoo_ehr ~ PC1_500m + (1|Site_ID), family = binomial, data = aa_join)
E3 <- glmer(zoo_ehr ~ PC1_500m + Field_Type + (1|Site_ID), family = binomial, data = aa_join)
E4 <- glmer(zoo_ehr ~ PC1_500m * Field_Type + (1|Site_ID), family = binomial, data = aa_join)

AICc(E0,E1,E2,E3,E4)
```

model with lowest AIC includes PC1 and field type, but no interaction

look at the model 
```{r}
summary(E3)
```

#### model presence/absence of Theileria cervi at the individual level

try out mixed-effects GLM w/ binomial distribution - using Site ID as a random effect
```{r, message=FALSE}
T0 <- glmer(Theileria_cervi ~ 1 + (1|Site_ID), family = binomial, data = aa_join)
T1 <- glmer(Theileria_cervi ~ Field_Type + (1|Site_ID), family = binomial, data = aa_join)
T2 <- glmer(Theileria_cervi ~ PC1_500m + (1|Site_ID), family = binomial, data = aa_join)
T3 <- glmer(Theileria_cervi ~ PC1_500m + Field_Type + (1|Site_ID), family = binomial, data = aa_join)
T4 <- glmer(Theileria_cervi ~ PC1_500m * Field_Type + (1|Site_ID), family = binomial, data = aa_join)

AICc(T0,T1,T2,T3,T4)
```

null model is better 

## try using GLMs for proportional data 

summarize proportion of infected AA with any zoonotic pathogen, ehrlichia, and theileria cervi by site.  
also add column indicating pres/abs of the pathogen by site 

all zoonotic pathogens
```{r, message=FALSE}
aa_zoo <- aa_join %>%
  group_by(Site_ID, zoo_path) %>%
  dplyr::summarise(n=n()) %>%
  pivot_wider(names_from = zoo_path, values_from = n)

colnames(aa_zoo) <- c("Site_ID", "neg", "pos")
aa_zoo$pos[is.na(aa_zoo$pos)] <- 0
aa_zoo$sampled <- aa_zoo$neg+aa_zoo$pos
aa_zoo$prop <- aa_zoo$pos/aa_zoo$sampled
aa_zoo$pos_site <- 0
aa_zoo$pos_site[aa_zoo$pos > 0] <- 1 
  
aa_zoo[17,] <- list("UG", 0, 0, 0, 0, 0)
aa_zoo[18,] <- list("GA", 0, 0, 0, 0, 0)

aa_zoo
```

zoonotic ehrlichia 
```{r}
aa_ehr <- aa_join %>%
  group_by(Site_ID, zoo_ehr) %>%
  dplyr::summarise(n=n()) %>%
  pivot_wider(names_from = zoo_ehr, values_from = n)

colnames(aa_ehr) <- c("Site_ID", "neg", "pos")
aa_ehr$pos[is.na(aa_ehr$pos)] <- 0
aa_ehr$sampled <- aa_ehr$neg+aa_ehr$pos
aa_ehr$prop <- aa_ehr$pos/aa_ehr$sampled
aa_ehr$pos_site <- 0
aa_ehr$pos_site[aa_ehr$pos > 0] <- 1 

aa_ehr[17,] <- list("UG", 0, 0, 0, 0, 0)
aa_ehr[18,] <- list("GA", 0, 0, 0, 0, 0)

aa_ehr
```

theileria cervi
```{r}
aa_tc <- aa_join %>%
  group_by(Site_ID, Theileria_cervi) %>%
  dplyr::summarise(n=n()) %>%
  pivot_wider(names_from = Theileria_cervi, values_from = n)

colnames(aa_tc) <- c("Site_ID", "neg", "pos")
aa_tc$pos[is.na(aa_tc$pos)] <- 0
aa_tc$sampled <- aa_tc$neg+aa_tc$pos
aa_tc$prop <- aa_tc$pos/aa_tc$sampled
aa_tc$pos_site <- 0
aa_tc$pos_site[aa_tc$pos > 0] <- 1 

aa_tc[17,] <- list("UG", 0, 0, 0, 0, 0)
aa_tc[18,] <- list("GA", 0, 0, 0, 0, 0)
  
aa_tc
```


### proportion of any zoonotic pathogen (full dataset)

join site covariates to aa_zoo summary 
```{r}
aa_zoo_join <- aa_zoo %>%
  left_join(covar, by = "Site_ID")
aa_zoo_join
```

plot the distribution of proportions
```{r}
hist(aa_zoo_join$prop)
```

try transforming the proportions 
let's go with a arcsine tranformation (the alternative logit transformation does something weird with 0's)
```{r}
asinTransform <- function(p) { asin(sqrt(p)) }

aa_zoo_join$prop_asin <- asinTransform(aa_zoo_join$prop)

hist(aa_zoo_join$prop_asin)
```


now try some glm's with original proportions 
```{r, message=FALSE}
Z0 <- glm(prop ~ 1, family = binomial, data = aa_zoo_join, weights = sampled)
Z1 <- glm(prop ~ Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)
Z2 <- glm(prop ~ PC1_500m, family = binomial, data = aa_zoo_join, weights = sampled)
Z3 <- glm(prop ~ PC1_500m + Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)
Z4 <- glm(prop ~ PC1_500m * Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```

all models are VERY similar... models with explanatory variables are no better than the null


now try some glm's with the transformed data 
```{r, message=FALSE}
Z0 <- glm(prop_asin ~ 1, family = binomial, data = aa_zoo_join, weights = sampled)
Z1 <- glm(prop_asin ~ Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)
Z2 <- glm(prop_asin ~ PC1_500m, family = binomial, data = aa_zoo_join, weights = sampled)
Z3 <- glm(prop_asin ~ PC1_500m + Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)
Z4 <- glm(prop_asin ~ PC1_500m * Field_Type, family = binomial, data = aa_zoo_join, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```

alrighty, the model with PC1 at 500m, field type, AND there interaction is the top model

look at the model 
```{r}
summary(Z4)
```

pseudo R2
```{r}
100*(78.028/60.377)/78.028
```
this is very low... maybe there is a different way to calculate a psuedo R2 for proportion data 

model validation 
```{r}
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(Z4, which = 1)
plot(Z4, which = 2)
plot(Z4, which = 3)
plot(Z4, which = 5)
```
sites 5, 9, and 15 are outside the envelope


### proportion of any zoonotic pathogen (screened dataset) 
repeat models for propotion but screen out sites 5,9, and 15 

```{r, message=FALSE}
aa_zoo_join_screen <- aa_zoo_join[-c(5,9,15),]

Z0.screen <- glm(prop_asin ~ 1, family = binomial, data = aa_zoo_join_screen, weights = sampled)
Z1.screen <- glm(prop_asin ~ Field_Type, family = binomial, data = aa_zoo_join_screen, weights = sampled)
Z2.screen <- glm(prop_asin ~ PC1_500m, family = binomial, data = aa_zoo_join_screen, weights = sampled)
Z3.screen  <- glm(prop_asin ~ PC1_500m + Field_Type, family = binomial, data = aa_zoo_join_screen, weights = sampled)
Z4.screen  <- glm(prop_asin ~ PC1_500m * Field_Type, family = binomial, data = aa_zoo_join_screen, weights = sampled)

AICc(Z0.screen , Z1.screen , Z2.screen , Z3.screen , Z4.screen )
```

lowest AIC is now the null model... 

```{r}
summary(Z1.screen)
summary(Z2.screen)
```

okay so the proportion of A. americanum infected with a zoonotic pathogen is NOT influenced by field type or the PC1 axis 


### proportion of any ehrlichia pathogen (full dataset)

join site covariates to aa_ehr summary 
```{r}
aa_ehr_join <- aa_ehr %>%
  left_join(covar, by = "Site_ID")
aa_ehr_join
```

plot the distribution of proportions
```{r}
hist(aa_ehr_join$prop)
```

try transforming the proportions 
let's go with a arcsine tranformation (the alternative logit transformation does something weird with 0's)
```{r}
asinTransform <- function(p) { asin(sqrt(p)) }

aa_ehr_join$prop_asin <- asinTransform(aa_ehr_join$prop)

hist(aa_ehr_join$prop_asin)
```


now try some glm's with original proportions 
```{r, message=FALSE}
Z0 <- glm(prop ~ 1, family = binomial, data = aa_ehr_join, weights = sampled)
Z1 <- glm(prop ~ Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)
Z2 <- glm(prop ~ PC1_500m, family = binomial, data = aa_ehr_join, weights = sampled)
Z3 <- glm(prop ~ PC1_500m + Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)
Z4 <- glm(prop ~ PC1_500m * Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```

model with PC1 and field type best 

```{r}
summary(Z3)
```

pseudo R2
```{r}
100*(17.173/10.615)/17.173
```

model validation 
```{r}
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(Z3, which = 1)
plot(Z3, which = 2)
plot(Z3, which = 3)
plot(Z3, which = 5)
```

probably do not need to filter out any sites... 

now try some glm's with the transformed data 
```{r, message=FALSE}
Z0 <- glm(prop_asin ~ 1, family = binomial, data = aa_ehr_join, weights = sampled)
Z1 <- glm(prop_asin ~ Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)
Z2 <- glm(prop_asin ~ PC1_500m, family = binomial, data = aa_ehr_join, weights = sampled)
Z3 <- glm(prop_asin ~ PC1_500m + Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)
Z4 <- glm(prop_asin ~ PC1_500m * Field_Type, family = binomial, data = aa_ehr_join, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```
model with PC1 and field type best 

```{r}
summary(Z3)
```


pseudo R2
```{r}
100*(69.617/47.661)/69.617
```
this is very low... maybe there is a different way to calculate a psuedo R2 for proportion data  
this is even lower than the untransformed data... put the p-values are more significant... 


model validation 
```{r}
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(Z3, which = 1)
plot(Z3, which = 2)
plot(Z3, which = 3)
plot(Z3, which = 5)
```

sites 2 and 15 may need be be removed... 


### proportion of any ehrlichia pathogen (screened dataset)

only the dataset with the transformed response might need sites removed 
```{r, message=FALSE}
aa_ehr_join_screen <- aa_ehr_join[-c(2,15),]

Z0 <- glm(prop_asin ~ 1, family = binomial, data = aa_ehr_join_screen, weights = sampled)
Z1 <- glm(prop_asin ~ Field_Type, family = binomial, data = aa_ehr_join_screen, weights = sampled)
Z2 <- glm(prop_asin ~ PC1_500m, family = binomial, data = aa_ehr_join_screen, weights = sampled)
Z3 <- glm(prop_asin ~ PC1_500m + Field_Type, family = binomial, data = aa_ehr_join_screen, weights = sampled)
Z4 <- glm(prop_asin ~ PC1_500m * Field_Type, family = binomial, data = aa_ehr_join_screen, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```
model with just field type best 

```{r}
summary(Z1)
```

field site is not significant in the best model 


### proportion of Theileria cervi (full dataset)

join site covariates to aa_tc summary 
```{r}
aa_tc_join <- aa_tc %>%
  left_join(covar, by = "Site_ID")
aa_tc_join
```

plot the distribution of proportions
```{r}
hist(aa_tc_join$prop)
```

try transforming the proportions 
let's go with a arcsine tranformation (the alternative logit transformation does something weird with 0's)
```{r}
asinTransform <- function(p) { asin(sqrt(p)) }

aa_tc_join$prop_asin <- asinTransform(aa_tc_join$prop)

hist(aa_tc_join$prop_asin)
```

transforming actually looks worse so i will only try original proportions 

now try some glm's with original proportions 
```{r}
Z0 <- glm(prop ~ 1, family = binomial, data = aa_tc_join, weights = sampled)
Z1 <- glm(prop ~ Field_Type, family = binomial, data = aa_tc_join, weights = sampled)
Z2 <- glm(prop ~ PC1_500m, family = binomial, data = aa_tc_join, weights = sampled)
Z3 <- glm(prop ~ PC1_500m + Field_Type, family = binomial, data = aa_tc_join, weights = sampled)
Z4 <- glm(prop ~ PC1_500m * Field_Type, family = binomial, data = aa_tc_join, weights = sampled)

AICc(Z0, Z1, Z2, Z3, Z4)
```

null model is best 

```{r}
summary(Z0)
```


## presence/absence at the site level

### presense/absence of any zoonotic pathogen at the site level

try out a set of models that is for pres/abs of pathogen at the site level 
```{r}
P0 <- glm(pos_site ~ 1, family = binomial, data = aa_zoo_join)
P1 <- glm(pos_site ~ Field_Type, family = binomial, data = aa_zoo_join)
P2 <- glm(pos_site ~ PC1_500m, family = binomial, data = aa_zoo_join)
P3 <- glm(pos_site ~ PC1_500m + Field_Type, family = binomial, data = aa_zoo_join)
P4 <- glm(pos_site ~ PC1_500m * Field_Type, family = binomial, data = aa_zoo_join)

AICc(P0, P1, P2, P3, P4)
```

top model for pres/abs at a site is P2 (PC1 at 500m) 

look at the model 
```{r}
summary(P2)
```

model fit 
```{r}
100*(24.953-19.194)/24.953
```


model validation 
```{r}
plot(P2, which = 1)
plot(P2, which = 2)
plot(P2, which = 3)
plot(P2, which = 4)
plot(P2, which = 5)
```


plot the prediction 

```{r}
mydata <- data.frame(PC1_500m = seq(-3,3, length = 50))
mydata$pred <- predict(P2, newdata = mydata, type = "response")
plot(x = aa_zoo_join$PC1_500m, y = aa_zoo_join$pos_site)
lines(mydata$PC1_500m, mydata$pred)
```


### presense/absence of any Ehrlichia pathogen at the site level

try out a set of models that is for pres/abs of pathogen at the site level 
```{r}
#aa_ehr_join$pos_site <- as.factor(aa_ehr_join$pos_site)
#levels(aa_ehr_join$pos_site) <- c("0", "1")

P0 <- glm(pos_site ~ 1, family = binomial, data = aa_ehr_join)
P1 <- glm(pos_site ~ Field_Type, family = binomial, data = aa_ehr_join)
P2 <- glm(pos_site ~ PC1_500m, family = binomial, data = aa_ehr_join)
P3 <- glm(pos_site ~ PC1_500m + Field_Type, family = binomial, data = aa_ehr_join)
P4 <- glm(pos_site ~ PC1_500m * Field_Type, family = binomial, data = aa_ehr_join)

AICc(P0, P1, P2, P3, P4)
```

top model for pres/abs at a site is P1 (field type) 

look at the model 
```{r}
summary(P1)
summary(P3)
```

model fit 
```{r}
100*(24.057-17.736)/24.057
100*(24.057-15.470 )/24.057
```

AIC and model fit is better for the model that include both field type and PC1 


model validation 
```{r}
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(P3, which = 1)
plot(P3, which = 2)
plot(P3, which = 3)
plot(P3, which = 5)
```

looks okay!

plot the prediction 
```{r}
mydata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Natural","Manicured"))
mydata$pred <- predict(P3, newdata = mydata, type = "response")

ggplot(aa_ehr_join, aes(x = PC1_500m, y = pos_site, group = Field_Type)) + 
  geom_point(aes(color = Field_Type)) + 
  geom_line(data = mydata, aes(x = PC1_500m, y = pred, color = Field_Type)) +
  theme_bw()
```

i f'd up changing the pres/ab to a factor and now it is funky... 



### presense/absence of Theileria cervi at the site level

try out a set of models that is for pres/abs of pathogen at the site level 
```{r}
P0 <- glm(pos_site ~ 1, family = binomial, data = aa_tc_join)
P1 <- glm(pos_site ~ Field_Type, family = binomial, data = aa_tc_join)
P2 <- glm(pos_site ~ PC1_500m, family = binomial, data = aa_tc_join)
P3 <- glm(pos_site ~ PC1_500m + Field_Type, family = binomial, data = aa_tc_join)
P4 <- glm(pos_site ~ PC1_500m * Field_Type, family = binomial, data = aa_tc_join)

AICc(P0, P1, P2, P3, P4)
```

top model for pres/abs at a site is P2 (PC1 at 500m) 

look at the model 
```{r}
summary(P2)
```

model fit 
```{r}
100*(24.953-12.990)/24.953
```


model validation 
```{r}
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(P2, which = 1)
plot(P2, which = 2)
plot(P2, which = 3)
plot(P2, which = 5)
```


plot the prediction 

```{r}
mydata <- data.frame(PC1_500m = seq(-3,3, length = 50))
mydata$pred <- predict(P2, newdata = mydata, type = "response")
plot(x = aa_tc_join$PC1_500m, y = aa_tc_join$pos_site)
lines(mydata$PC1_500m, mydata$pred)
```
