---
title: "PCA of landcover composition in Alachua Co. Florida"
author: "Kimberly Ledger"
date: "3/3/2022"
output: github_document
---

this code runs a principal components analysis (PCA) using land cover composition proportions using a 1km buffer around potential tick collection sites in Alachua County, Florida 

last update: 20 May 2022

load libraries 
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(ggbiplot)
```


read in the data
```{r}
lc <- read.csv("AlCo_landcover_allscales_sort.csv")
head(lc)
```

this is the landcover classification

*  1 = Agriculture 
*  2 = Commercial
*  3 = Forest
*  4 = Lawn 
*  5 = Residential
*  6 = Water
*  7 = Wetland

sampling site selection was based on PC1 using land cover proportions using a 1km buffer around all potential sites 

filter data to only include land cover at 1km 
```{r}
lc_1km <- lc[,c(26:32)]
head(lc_1km)
```

log transform and apply z-score standardization to the data 
```{r}
lc_1km_log <- log1p(lc_1km)   
z_lc <- lc_1km_log
```

update column names 
```{r}
colnames(z_lc) <- c("agriculture", "commercial", "forest", "lawn", "residential", "water", "wetland")
```

run a PCA using the function princomp from the stats package
```{r}
lc_pca <- princomp(z_lc, cor=F)
#cor=F, because we are using the covariance matrix instead of the correlation matrix 

summary(lc_pca)

#get factor loadings for each PC
loadings(lc_pca)

# get factor scores for each site 
lc_pca$scores

#access using a scree plot 
plot(lc_pca, type="lines")
```

plot the first two axes of the PCA - landcovers
```{r}
plot(lc_pca$loadings,type="n",xlab="PC 1", ylab="PC 2")
text(lc_pca$loadings, labels=as.character(colnames(z_lc)), pos=1, cex=1)
```

plot the first two axes of the PCA - sites
```{r}
plot(lc_pca$scores,type="n",xlab="PC 1", ylab="PC 2")
text(lc_pca$scores, labels=as.character(rownames(z_lc)), pos=1, cex=1)
```

biplot 
```{r}
biplot(lc_pca$scores,lc_pca$loading,xlab="PC 1", ylab="PC 2",ylim=c(-0.2,0.4), xlim=c(-0.4,0.5))
```

biplot - replace site numbers with '*'
```{r}
biplot(lc_pca$scores,lc_pca$loading,xlabs= rep("*",52), xlab="PC 1", ylab="PC 2",ylim=c(-0.2,0.4), xlim=c(-0.4,0.5))

```

now try ggbiplot 

```{r}
lc <- lc %>%
  unite("GS", Field_Type, sampled, sep = "/", remove = F)

unique(lc$GS)
lc.newgroups <- as.factor(lc$GS)
lc.new <- factor(lc.newgroups, levels = c("Manicured/Yes", "Natural/Yes", "Manicured/No", "Natural/No"))


biplot <- ggbiplot(lc_pca, groups = lc.new) +
  scale_color_manual(name = "Habitat Type/Sampled?", values = c("#440154FF", "#21908CFF", "#efb3fe", "#acecea")) + 
  theme_classic() + 
  xlim(c(-2,2))

biplot
```

ggsave("Figures/biplot.png", dpi=300, height = 6, width = 8)

the issue here is that site points are overlapping... but i rearranged sites so that all sampled sites plot on top 

negative values of PC1 are associated with have high proportions of commercial and residential and the positive values of PC1 are associated with high proportions of forest cover (and to some extent wetland)
