---
title: "tick_pathogens"
author: "Kimberly Ledger"
date: "4/18/2022"
output: github_document
---

Code for analyses of pathogens in Alachua County ticks.  
Updated 20 May 2022

load libraries
```{r, message=FALSE}
library(tidyverse) # for data manipulation
library(ggplot2) # for visualizations
library(lme4) #for glm()
library(MuMIn) # for AICc() 
library(stargazer) # for stargazer() model comparison
library(MASS) # for glm.nb()
```


# Part 1: Prepare the data
Read in raw Alachua County tick data
```{r}
ticks <- read.csv("AlachuaTicksAllSites.csv")
str(ticks)
```

Read in raw pathogen data 
```{r}
pathogen <- read.csv("AlCo_Pathogens.csv")
str(pathogen)
```


Join pathogen results with the raw tick data by the gDNA_Sample and Collection_Sample identities
```{r}
data <- ticks %>%
  dplyr::left_join(pathogen, by = c("gDNA_Sample", "Collection_Sample")) %>%
  mutate_all(na_if, "")

head(data)
```

note: there are NA's in the data for any tick sample that was not processed in the lab

Read in the covariate data for all 18 sites 
This site contains information on if the environment type was manicured or natural, and the forest, PC1 and PC2 landscape buffers
```{r}
covar <- read.csv("AlCo_SitesCovariates.csv")  ## 18 sites 
head(covar)
```

# Part 2: Create summary table with response variables 

* = presence/absence of infected ticks by site 
* = abundance of infected ticks per site 
* = richness of pathogens per site

important: we are only considering microorganisms that are known to be human or veterinary important pathogens as "pathogens" in these analyses 

```{r}
pathogen_df <- data %>%
  filter(!is.na(gDNA_Sample)) %>%
  group_by(Site_ID) %>%
  dplyr::summarise(Babesia_sp_Coco = sum(Babesia_sp_Coco),
            Babesia_odocoieli = sum(Babesia_odocoieli),
            Cytauxozoon_felis = sum(Cytauxozoon_felis),
            Ehrlichia_chaffeensis = sum(Ehrlichia_chaffeensis),
            Ehrlichia_ewingii = sum(Ehrlichia_ewingii),
            Ehrlichia_sp_PanolaMountain = sum(Ehrlichia_sp_PanolaMountain),
            Rickettsia_parkeri = sum(Rickettsia_parkeri), 
            sum_path = sum(Babesia_sp_Coco, Babesia_odocoieli, Cytauxozoon_felis,
                           Ehrlichia_chaffeensis, Ehrlichia_ewingii, Ehrlichia_sp_PanolaMountain,
                           Rickettsia_parkeri))
```

add in row for site with no ticks
```{r}
pathogen_df[18,] <- list("GA", 0, 0, 0, 0, 0, 0, 0, 0)
```


create df for pathogen richness 
```{r}
richness_df <- pathogen_df

richness_df$Babesia_sp_Coco[richness_df$Babesia_sp_Coco > 0] <- 1 
richness_df$Babesia_odocoieli[richness_df$Babesia_odocoieli > 0] <- 1 
richness_df$Cytauxozoon_felis[richness_df$Cytauxozoon_felis > 0] <- 1 
richness_df$Ehrlichia_chaffeensis[richness_df$Ehrlichia_chaffeensis > 0] <- 1 
richness_df$Ehrlichia_ewingii[richness_df$Ehrlichia_ewingii > 0] <- 1 
richness_df$Ehrlichia_sp_PanolaMountain[richness_df$Ehrlichia_sp_PanolaMountain > 0] <- 1
richness_df$Rickettsia_parkeri[richness_df$Rickettsia_parkeri > 0] <- 1 
richness_df$Babesia_sp_Coco[richness_df$Babesia_sp_Coco > 0] <- 1 

richness_df <- richness_df %>%
  dplyr::select(!sum_path) %>%
  dplyr::mutate(all_rich = rowSums(across(where(is.numeric))))
```

put the explanitory (field_type and PC1 @ 500m) and response variables into one data frame 
```{r}
covar1 <- covar %>%
  dplyr::select(Site_ID, Field_Type, PC1_500m)

in_abun <- pathogen_df %>%
  dplyr::select(Site_ID, sum_path)

in_rich <- richness_df %>%
  dplyr::select(Site_ID, all_rich)

join_path <- covar1 %>%
  left_join(in_abun) %>%
  left_join(in_rich)

join_path$pres <- 0 
join_path$pres[join_path$all_rich > 0] <- 1  ## set coinfections to 1 

join_path <- join_path %>%
  arrange(PC1_500m)
```


# Part 3: model time! 

## presence of pathogens at the site level 
## (i.e. whether or not at least one infected tick was found with any pathogen in a site)

binomial glm's 
```{r}
P0 <- glm(pres ~ 1, family = binomial, data = join_path)
P1 <- glm(pres ~ Field_Type, family = binomial, data = join_path)
P2 <- glm(pres ~ PC1_500m, family = binomial, data = join_path)
P3 <- glm(pres ~ PC1_500m + Field_Type, family = binomial, data = join_path)

AICc(P0, P1, P2, P3)
```

compare the top models  
```{r}
stargazer(P2, P3, type = "text",
          intercept.bottom = FALSE, 
          single.row=FALSE,     
          notes.append = FALSE, 
          header=FALSE)
```


top model summary and model validation plots
```{r}
summary(P3)

par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(P3, which = 1)
plot(P3, which = 2)
plot(P3, which = 3)
#plot(P3, which = 4)
plot(P3, which = 5)
```

sites 5 and 12 could be are problematic... consider removing later


pretty plot of the predictions - start by making model predictions
```{r}
## grad the inverse link function
ilink <- family(P3)$linkinv

ndata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Manicured", "Natural"))
ndata$pres <- predict(P3, newdata = ndata, type = "response")
ndata <- bind_cols(ndata, setNames(as_tibble(predict(P3, ndata, se.fit = TRUE)[1:2]),
                                   c('fit_link','se_link')))
## create the interval and backtransform
ndata <- mutate(ndata,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)))
## show
ndata
```

plot
```{r pres_infected}
library(viridisLite)
mycol <- viridisLite::viridis(3)

v1 <- ggplot(data = join_path, aes(x = PC1_500m, y = pres, fill = Field_Type)) + 
  geom_point(cex = 4, alpha = 0.7, color = "black", pch = 21, position = position_jitter(w = 0.05, h = 0)) +
  scale_fill_manual(values = mycol) + 
  geom_line(data = ndata, aes(color = Field_Type), size = 1.5) +
  #geom_ribbon(data = ndata,
  #                aes(ymin = right_lwr, ymax = right_upr, fill=Field_Type),
  #                alpha = 0.1) + 
  scale_color_manual(values = mycol) + 
  theme(#axis.text.y= element_blank(),   
        #axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none") + 
  #scale_color_discrete(name = "Environment Type", labels = c("manicured", "natural")) + 
  ylab("Probability of Infected Tick") +
  xlab("Development Gradient")
v1
```



remove problematic sites 

```{r}
join_path_screen <- join_path[-c(5,12),]
```


binomial glm's 
```{r}
P0 <- glm(pres ~ 1, family = binomial, data = join_path_screen)
P1 <- glm(pres ~ Field_Type, family = binomial, data = join_path_screen)
P2 <- glm(pres ~ PC1_500m, family = binomial, data = join_path_screen)
P3 <- glm(pres ~ PC1_500m + Field_Type, family = binomial, data = join_path_screen)

AICc(P0, P1, P2, P3)
```

compare the top models  
```{r}
stargazer(P2, P3, type = "text",
          intercept.bottom = FALSE, 
          single.row=FALSE,     
          notes.append = FALSE, 
          header=FALSE)
```

models do not produce warning of "Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred"
i use models that retain all sites to keep the sample size as large as possible... 




## abundance of infected ticks per site

```{r}
hist(join_path$sum_path)
```

try a negative binomial glm 
```{r}
A0 <- glm.nb(sum_path ~ 1, data = join_path)
A1 <- glm.nb(sum_path ~ Field_Type, data = join_path)
A2 <- glm.nb(sum_path ~ PC1_500m, data = join_path)
A3 <- glm.nb(sum_path ~ PC1_500m + Field_Type, data = join_path)
A4 <- glm.nb(sum_path ~ PC1_500m * Field_Type, data = join_path)

AICc(A0, A1, A2, A3, A4)
```

model summary and validation
```{r}
summary(A3)
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(A3, which = 1)
plot(A3, which = 2)
plot(A3, which = 3)
#plot(A3, which = 4)
plot(A3, which = 5)
```

site 11 could be problematic... 

pretty plot of the predictions 
```{r}
## grad the inverse link function
ilink <- family(A3)$linkinv
mydata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Manicured", "Natural"))
mydata$sum_path <- predict(A3, newdata = mydata, type = "response")
mydata <- bind_cols(mydata, setNames(as_tibble(predict(A3, mydata, se.fit = TRUE)[1:2]),
                                   c('fit_link','se_link')))
## create the interval and backtransform
mydata <- mutate(mydata,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)))
## show
mydata
```

```{r abund_infected}
v2 <- ggplot(join_path, aes(x = PC1_500m, y = sum_path, group = Field_Type)) + 
  geom_point(aes(fill = Field_Type), cex = 4, alpha = 0.7, 
             color = "black", pch = 21) +
  scale_fill_manual(values = mycol, name = "Habitat Type", labels = c("manicured", "natural")) + 
  geom_line(data = mydata, aes(color = Field_Type)) +
  geom_ribbon(data = mydata,
                  aes(ymin = right_lwr, ymax = right_upr, fill=Field_Type),
                  alpha = 0.2) + 
  scale_color_manual(values = mycol, name = "Habitat Type", labels = c("manicured", "natural")) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = c(0.2,0.6)) + 
  ylab("Abundance of Infected Ticks") +
  xlab("Development Gradient")
v2
```



An alternative would be to log-transform data and run linear model... 

```{r}
join_path$log_sum_path <- log1p(join_path$sum_path)
join_path$log_all_rich <- log1p(join_path$all_rich)
```


try lms 
```{r}
L0 <- lm(log_sum_path ~ 1, data = join_path)
L1 <- lm(log_sum_path ~ Field_Type, data = join_path)
L2 <- lm(log_sum_path ~ PC1_500m, data = join_path)
L3 <- lm(log_sum_path ~ PC1_500m + Field_Type, data = join_path)
L4 <- lm(log_sum_path ~ PC1_500m * Field_Type, data = join_path)

AICc(L0, L1, L2, L3, L4)
```

model summary and validation
```{r}
summary(L3)
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(L3, which = 1)
plot(L3, which = 2)
plot(L3, which = 3)
#plot(L3, which = 4)
plot(L3, which = 5)
```

pretty plot of the predictions 
```{r}
mydata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Manicured", "Natural"))
mydata$pred <- predict(L3, newdata = mydata, type = "response")

v3 <- ggplot(join_path, aes(x = PC1_500m, y = log_sum_path, group = Field_Type)) + 
  geom_point(aes(fill = Field_Type, cex = 1.5, alpha = 0.5), 
             color = "black", pch = 21) +
  scale_fill_manual(values = mycol) + 
  geom_line(data = mydata, aes(x = PC1_500m, y = pred, color = Field_Type)) +
  scale_color_manual(values = mycol) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none") + 
  #scale_color_discrete(name = "Environment Type", labels = c("manicured", "natural")) + 
  ylab("log(Abundance of Infected Ticks)") +
  xlab("Development Gradient")
v3
```

screen data to remove site 11

```{r}
join_path_screen <- join_path[-11,]
```


try a negative binomial glm 
```{r}
A0 <- glm.nb(sum_path ~ 1, data = join_path_screen)
A1 <- glm.nb(sum_path ~ Field_Type, data = join_path_screen)
A2 <- glm.nb(sum_path ~ PC1_500m, data = join_path_screen)
A3 <- glm.nb(sum_path ~ PC1_500m + Field_Type, data = join_path_screen)
A4 <- glm.nb(sum_path ~ PC1_500m * Field_Type, data = join_path_screen)

AICc(A0, A1, A2, A3, A4)
```

model summary and validation
```{r}
summary(A3)
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(A3, which = 1)
plot(A3, which = 2)
plot(A3, which = 3)
#plot(A3, which = 4)
plot(A3, which = 5)
```

removing site 11 seems to do more harm than good...



## pathogen richness per site
```{r}
hist(join_path$all_rich)
```

try a negative binomial glm 
```{r}
R0 <- glm.nb(all_rich ~ 1, data = join_path)
R1 <- glm.nb(all_rich ~ Field_Type, data = join_path)
R2 <- glm.nb(all_rich ~ PC1_500m, data = join_path)
R3 <- glm.nb(all_rich ~ PC1_500m + Field_Type, data = join_path)
R4 <- glm.nb(all_rich ~ PC1_500m * Field_Type, data = join_path)

AICc(R0, R1, R2, R3, R4)
```

model summary and validation
```{r}
summary(R3)
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(R3, which = 1)
plot(R3, which = 2)
plot(R3, which = 3)
#plot(R3, which = 4)
plot(R3, which = 5)
```


pretty plot of the predictions 
```{r}
ilink <- family(R3)$linkinv

mydata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Manicured", "Natural"))
mydata$all_rich <- predict(R3, newdata = mydata, type = "response")

mydata <- bind_cols(mydata, setNames(as_tibble(predict(R3, ndata, se.fit = TRUE)[1:2]),
                                   c('fit_link','se_link')))

## create the interval and backtransform
mydata <- mutate(mydata,
                fit_resp  = ilink(fit_link),
                right_upr = ilink(fit_link + (2 * se_link)),
                right_lwr = ilink(fit_link - (2 * se_link)))
## show
mydata
```

```{r pathogen_richness}
v4 <- ggplot(join_path, aes(x = PC1_500m, y = all_rich, group = Field_Type)) + 
  geom_point(aes(fill = Field_Type), cex = 4, alpha = 0.7, 
             color = "black", pch = 21) +
  scale_fill_manual(values = mycol) + 
  geom_line(data = mydata, aes(color = Field_Type)) +
  geom_ribbon(data = mydata,
                  aes(ymin = right_lwr, ymax = right_upr, fill=Field_Type),
                  alpha = 0.2) + 
  scale_color_manual(values = mycol) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none") + 
  ylab("Pathogen Richness") +
  xlab("Development Gradient")
v4
```



or does it makes sense to log-transform data and run linear models? 

try lms 
```{r}
RL0 <- lm(log_all_rich ~ 1, data = join_path)
RL1 <- lm(log_all_rich ~ Field_Type, data = join_path)
RL2 <- lm(log_all_rich ~ PC1_500m, data = join_path)
RL3 <- lm(log_all_rich ~ PC1_500m + Field_Type, data = join_path)
RL4 <- lm(log_all_rich ~ PC1_500m * Field_Type, data = join_path)

AICc(RL0, RL1, RL2, RL3, RL4)
```

model summary and validation
```{r}
summary(RL3)
par(mfrow=c(2,2), mar=c(4,4,2,1))
plot(RL3, which = 1)
plot(RL3, which = 2)
plot(RL3, which = 3)
#plot(RL3, which = 4)
plot(RL3, which = 5)
```

pretty plot of the predictions 
```{r}
mydata <- expand.grid(PC1_500m = seq(-3,3, length = 50), Field_Type = c("Manicured", "Natural"))
mydata$pred <- predict(RL3, newdata = mydata, type = "response")

v5 <- ggplot(join_path, aes(x = PC1_500m, y = log_all_rich, group = Field_Type)) + 
  geom_point(aes(fill = Field_Type, cex = 1.5, alpha = 0.5), 
             color = "black", pch = 21) +
  scale_fill_manual(values = mycol) + 
  geom_line(data = mydata, aes(x = PC1_500m, y = pred, color = Field_Type)) +
  scale_color_manual(values = mycol) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none") + 
  #scale_color_discrete(name = "Environment Type", labels = c("manicured", "natural")) + 
  ylab("log(Pathogen Richness)") +
  xlab("Development Gradient")
v5
```

make final composite figure for manuscript 

```{r pathogen_fig, fig.width = 5, fig.height = 10}
library(cowplot)

plot_grid(v1, v2, v4, labels= "AUTO", ncol=1)
```



